\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[hidelinks]{hyperref}

\title{Emergent Gravity from the Gauged Hopf Lattice\\
\large Weak-Field Limit, $G_N$ Schema, and Lattice-to-Metric Bridge}
\author{Aaron Kinder \\ Independent Researcher}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
We record the continuum gravity scaffold of the gauged Hopf lattice: effective
stress-energy from holonomy and twist gradients, Einstein--Hilbert structure with
Newton constant fixed by frozen locks via
$G_N\sim 8\pi\lambda\Delta\omega/(\kappa W_g^2)$, and the Newtonian weak-field limit.
Gate~GR-1 enforces structural consistency; Gate~GR-2 checks analytic GR targets
with matched $G$. No claim of full nonlinear GR equivalence or new observational
discovery is made. Pre-merger GW freeze remains independent.
\end{abstract}

\section{Discipline and locks}

Frozen:
\[
W_g=\frac{350}{\pi},\qquad \kappa\approx 0.85,\qquad \phi_b^\star\approx 0.8145.
\]
Unified action~\cite{lag,rel} includes
$S_{\mathrm{grav}}= (16\pi G_N)^{-1}\int\sqrt{-g}\,R+\cdots$.
Independent GW pre-merger pipelines are not reopened by this scaffold.

\section{$G_N$ schema and tight SI bridge}

\begin{equation}
G_{\mathrm{schema}}
=\frac{8\pi\,\lambda\,\Delta\omega\,f(\langle\Theta\rangle)}{\kappa\,W_g^{2}}.
\label{eq:GN-schema}
\end{equation}
Continuum SI map (Gate~GR-3):
\begin{equation}
G_{\mathrm{SI}}
= G_{\mathrm{schema}}\times\frac{\hbar c}{m_\star^{2}},
\qquad
m_\star^{2}
= G_{\mathrm{schema}}(\mathrm{default})\times\frac{\hbar c}{G_{\mathrm{CODATA}}}.
\label{eq:SI-bridge}
\end{equation}
Thus $G_{\mathrm{SI}}(\mathrm{default})\equiv G_{\mathrm{CODATA}}$ with explicit invertible
$m_\star$; drifts in $\lambda,\Delta\omega,f$ (locks fixed) scale $G_{\mathrm{SI}}$
proportionally. Round-trip schema$\leftrightarrow$SI is checked numerically.

\section{Effective stress-energy}

From the free-energy density of the twist/holonomy sector,
\begin{equation}
\rho_{\mathrm{eff}}
\sim \frac{D}{8}|\nabla\Theta|^{2}
+\frac{\kappa}{2}\bar\Theta^{2},
\label{eq:rho-eff}
\end{equation}
with isotropic pressure scaffold $p\sim\tfrac13\rho_{\mathrm{grad}}$ for the gradient piece.
Trace $T=\rho-3p$ sources an effective curvature scalar
\begin{equation}
R_{\mathrm{eff}}\sim 8\pi G_{\mathrm{schema}}\,T.
\label{eq:R-eff}
\end{equation}

\section{Einstein limit (structure)}

\begin{equation}
G_{\mu\nu}+\Lambda g_{\mu\nu}=8\pi G_N T_{\mu\nu},
\label{eq:einstein}
\end{equation}
with $T_{\mu\nu}$ identified with holonomy/twist continuum stress (defects, bursts as
localized sources). Full metric PDE solutions are later work.

\section{Newtonian / weak-field limit}

\begin{equation}
\nabla^{2}\Phi=4\pi G\rho,
\qquad
\Phi=-\frac{GM}{r}.
\label{eq:newton}
\end{equation}
Linearized metric:
$g_{00}\approx -(1+2\Phi)$, $g_{ij}\approx (1-2\Phi)\delta_{ij}$ ($c=1$).

\section{Analytic precision targets (Gate~GR-2)}

With matched $G$: solar light deflection $\delta=4GM_\odot/(c^{2}R_\odot)\approx 1.75''$;
Mercury perihelion $\approx 43''$/cy; Shapiro scale $2GM_\odot/c^{3}\approx 9.85\,\mu\mathrm{s}$.
These validate the weak-field module structure, not a new ephemeris campaign.

\section{Lattice $\to$ metric Poisson PDE (Gate~GR-3)}

On a periodic 2-torus, holonomy defects in $\Theta$ source
$\rho_{\mathrm{eff}}$; solve
\begin{equation}
\nabla^{2}\Phi = 4\pi G\rho_{\mathrm{eff}}
\label{eq:poisson-lattice}
\end{equation}
by FFT (spectral residual). Weak-field metric
$g_{00}=-(1+2\Phi/c^{2})$. Pass requires tiny residual, attractive $\Phi$,
and $\mathrm{corr}(\rho,\Phi)<0$.

\section{Gates}

\begin{center}
\begin{tabular}{@{}llp{0.4\textwidth}@{}}
\hline
Gate & Status & Role \\
\hline
GR-1 & PASS (scaffold) & $G_N$ schema, $\rho_{\mathrm{eff}}$, Newton limit, locks \\
GR-2 & PASS (structure) & Analytic GR targets with matched $G$ \\
GR-3 & PASS (scaffold) & SI bridge $m_\star$ + lattice$\to$metric PDE residual \\
\hline
\end{tabular}
\end{center}

Companion software (optional reproducibility) implements GR-1/2/3 checks.

\section{What is not claimed}

Full nonlinear Einstein equivalence, cosmology, or SI $G_N$ without continuum matching.
FAIL demotes the gravity mapping, not the core locks.

\paragraph{Series note.}
This note is Paper~IV of the gated scaffold series (Papers~I--IV).

\begin{thebibliography}{9}
\bibitem{lag}
A.~Kinder, \emph{Unified Action Principle for the Gauged Hopf Lattice},
companion note (Paper~I), Jul.~2026.
\bibitem{rel}
A.~Kinder, \emph{Relativistic Completion of the Gauged Hopf Lattice},
companion note (Paper~II), Jul.~2026.
\bibitem{sm}
A.~Kinder, \emph{SM Particle Content from Gauged Hopf Lattice Modes},
companion note (Paper~III), Jul.~2026.
\end{thebibliography}

\end{document}
