\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}

\title{Unified Action Principle for the Gauged Hopf Lattice\\
\large with SU(3)$\times$SU(2)$\times$U(1) Scaffolding\\
\normalsize (Phase~1.1 --- invariant\_hunt extension)}
\author{Aaron Kinder \\ Independent Researcher}
\date{April 2026 \\ \small Phase~1.1 extension: July 2026}

\begin{document}
\maketitle

\begin{abstract}
The conduit PDE of the gauged two-gyro Hopf lattice is a dissipative first-order
gradient flow and cannot be derived from a standard time-reversible Lagrangian alone.
Its clean variational structure is the Dirichlet (harmonic-map) energy on the
quaternion field $q(x,t)\in S^3$, projected to the scalar twist density
$\Theta=2\arccos(\operatorname{Re} q)$, plus explicit non-variational forcing for
the two-gyro drive, global gauge holonomy, and a potential for topological bursts.
We record the free-energy functional, its $L^2$-gradient flow, the reduction to the
scalar nonlinear twist PDE, and the conservative wave-map Lagrangian before the
over-damped limit.

\textbf{Phase~1.1.} We extend this foundation to a unified action with explicit
Yang--Mills kinetics for $G=\mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)$,
a Hopf topological density, and holonomy/braiding terms tied to the
\emph{frozen} locks $W_g=350/\pi$, $\kappa\approx 0.85$, $\phi_b^\star\approx 0.8145$.
Gate~A-P requires healthy kinetic signs (no ghosts), conduit reduction, and
$W_g$ stability under holonomy/gauge jitter. No claim is made that the SM spectrum
or Einstein gravity have been derived; those are later-phase gated deliverables.
\end{abstract}

\section{Underlying field: $q:T^3\times\mathbb{R}\to S^3$}

The microscopic field is a unit quaternion
\[
q=w+ix+jy+kz,\qquad w^2+x^2+y^2+z^2=1.
\]
Local twist density:
\[
\Theta(x,t)=2\arccos(\operatorname{Re} q)=2\arccos(w).
\]
Dirichlet energy (harmonic-map action):
\[
F[q]=\frac{D}{2}\int_{T^3}|\nabla q|^2\,\mathrm{d}^3x
\quad(|q|=1).
\]
The $L^2$-gradient flow on $S^3$ is the projected Laplacian (harmonic-map heat flow).

\section{Reduction to the scalar twist PDE}

When the rotation axis is aligned with the fixed gauge axis (global pointer),
$q\approx\cos\phi+\sin\phi\,k$ with $\phi=\Theta/2$ yields
\[
F[\Theta]\approx\frac{D}{8}\int_{T^3}|\nabla\Theta|^2\,\mathrm{d}^3x.
\]
The geometric Jacobian from the $S^3$ metric supplies the nonlinear term
$\tfrac{D}{2}\cot(\Theta/2)\,|\nabla\Theta|^2$. The conduit PDE is the effective
scalar equation obtained by projecting the quaternion harmonic-map flow.

\section{Full free-energy functional}

\begin{equation}
F[\Theta]
=\int_{T^3}\Biggl[
\frac{D}{8}|\nabla\Theta|^2
+U(\Theta)
+\frac{\kappa}{2}\bar\Theta^2
-\Delta\omega\,\Theta
\Biggr]\,\mathrm{d}^3x,
\label{eq:free-energy}
\end{equation}
with burst potential
\[
U'(\Theta)=-B(\Theta)=-C(\Theta-\theta_{\mathrm{crit}})_+^{p},
\qquad C\gg 1,\; p>1,
\]
and $\theta_{\mathrm{crit}}=\pi(1+\kappa)$ \cite{burst}.
$L^2$-gradient flow $\partial_t\Theta=-\delta F/\delta\Theta$ recovers
\begin{equation}
\partial_t\Theta
=D\Delta\Theta
+\tfrac{D}{2}\cot(\Theta/2)\,|\nabla\Theta|^2
+\Delta\omega
-\kappa\bar\Theta
+B(\Theta).
\label{eq:conduit}
\end{equation}

\section{Conservative Lagrangian (before over-damping)}

\begin{equation}
L[q]=\int_{T^3}\Biggl(
\frac{1}{2}|\partial_t q|^2
-\frac{D}{2}|\nabla q|^2
\Biggr)\,\mathrm{d}^3x
\quad(|q|=1),
\label{eq:wave-map}
\end{equation}
with Euler--Lagrange wave-map equation. Infinite friction recovers~\eqref{eq:conduit}.
Drive and gauge enter as non-variational external torques, or via a time-dependent
potential / global Lagrange multiplier for the mean twist.

\section{Phase~1.1: Unified action with SM gauge scaffold}

Define $U\in\mathrm{SU}(2)$ equivalent to $q\in S^3$. The Phase~1.1 action is
\begin{equation}
S=S_\sigma+S_{\mathrm{Skyrme}}+S_{\mathrm{YM}}
+S_{\mathrm{Hopf}}+S_{\mathrm{hol}}+S_{\mathrm{drive}}
+S_{\mathrm{burst}}+S_{\mathrm{grav}},
\label{eq:unified-action}
\end{equation}
with sector densities (symbolic module \texttt{src/action\_principle.py}):

\begin{align}
S_\sigma
&=\frac{1}{2\lambda^2}\int\mathrm{d}^4x\,\sqrt{-g}\,
\varepsilon_{\mu\nu}\,\mathrm{Tr}\!\bigl(D^\mu U\,D^\nu U^\dagger\bigr),
\label{eq:sigma}
\\
S_{\mathrm{Skyrme}}
&=\frac{1}{32\pi^2 e_S^2}\int\mathrm{d}^4x\,\sqrt{-g}\,
\mathrm{Tr}\!\bigl([D_\mu U U^\dagger,\,D_\nu U U^\dagger]^2\bigr),
\label{eq:skyrme}
\\
S_{\mathrm{YM}}
&=-\sum_{G\in\{\mathrm{SU}(3),\mathrm{SU}(2),\mathrm{U}(1)\}}
\frac{1}{4g_G^2}\int\mathrm{d}^4x\,\sqrt{-g}\,
F_{\mu\nu}^{(G)}F^{(G)\mu\nu},
\label{eq:ym}
\\
S_{\mathrm{Hopf}}
&=c_H\int H_{\mathrm{Hopf}}[q],
\qquad \int H_{\mathrm{Hopf}}\in\mathbb{Z},
\label{eq:hopf}
\\
S_{\mathrm{hol}}
&=-\frac{\kappa}{2}\int\mathrm{d}^4x\,\sqrt{-g}\,\bar\Theta^2
-\frac{W_g}{2}\int\mathrm{d}^4x\,\sqrt{-g}\,(\phi_b-\phi_b^\star)^2,
\label{eq:hol}
\\
S_{\mathrm{drive}}
&=-\Delta\omega\int\mathrm{d}^4x\,\sqrt{-g}\,\Theta,
\label{eq:drive}
\\
S_{\mathrm{burst}}
&=-\int\mathrm{d}^4x\,\sqrt{-g}\,U(\Theta),
\label{eq:burst-S}
\\
S_{\mathrm{grav}}
&=\frac{1}{16\pi G_N}\int\mathrm{d}^4x\,\sqrt{-g}\,R+\cdots
\quad\text{(Phase~3 target)}.
\label{eq:grav}
\end{align}

\paragraph{Frozen locks (inputs, not fits).}
\[
W_g=\frac{350}{\pi},\qquad
\kappa\approx 0.85,\qquad
\phi_b^\star\approx 0.8145,\qquad
\theta_{\mathrm{crit}}=\pi(1+\kappa).
\]
Hopf linking saturation
$\Theta_{\mathrm{link}}=2\pi W_g/(2W_g+1)\approx\pi$ remains geometric
\cite{burst}.

\paragraph{Scope discipline.}
Equation~\eqref{eq:ym} is a \emph{scaffold}: it places the correct gauge kinetics
and healthy signs. Mapping lattice excitations to the full SM spectrum (generations,
Yukawa, anomalies) is Phase~2. Emergent Einstein limit and precision tests are Phase~3.
No universal claim is made at Phase~1.1.

\section{Gate A-P (action-principle scrutiny)}

Any new gauge/Hopf term must pass:

\begin{enumerate}
\item \textbf{No ghosts.} Kinetic coefficients and holonomy Hessians have healthy signs:
  $D>0$, $g_i>0$, $e_S>0$, $\kappa>0$, $W_g>0$; wave-map time kinetic positive.
\item \textbf{Conduit reduction.} Over-damped / free-energy limit recovers~\eqref{eq:conduit}
  structure (drive, holonomy, Dirichlet, burst).
\item \textbf{$W_g$ stability.} Under jitter of $\kappa$, $\phi_b^\star$, and $g_i$
  (with $W_g$ held locked), positional residual $|W_g-350/\pi|$ remains zero and
  ghost checks remain healthy.
\item \textbf{Analytic consistency with GW pipelines.} New terms must not alter
  locked pre-merger template inputs ($W_g$, $\kappa$, $\phi_b$) used by
  \texttt{premerger\_core\_predict.py}.
\end{enumerate}

Executable check:
\begin{verbatim}
python scripts/action_principle_check.py
python scripts/action_principle_check.py --pde-smoke
\end{verbatim}

\section{Variational status table}

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Term in PDE / action & Origin & Variational? \\
\midrule
$D\Delta\Theta+\tfrac{D}{2}\cot(\Theta/2)|\nabla\Theta|^2$
  & Dirichlet / harmonic map on $S^3$ & Yes \\
$+\Delta\omega$ & Two-gyro drive mismatch & No (forcing) \\
$-\kappa\bar\Theta$ & Mean-field gauge holonomy & Yes \\
$+B(\Theta)$ & Burst potential $U(\Theta)$ & Yes \\
$S_{\mathrm{YM}}$ & Yang--Mills kinetics & Yes \\
$S_{\mathrm{Hopf}}$ & Topological Hopf density & Yes \\
$W_g(\phi_b-\phi_b^\star)^2$ pin & Braiding attractor & Yes \\
$S_{\mathrm{grav}}$ & Einstein--Hilbert (later) & Yes \\
\bottomrule
\end{tabular}
\caption{Origin and variational status of each term (extends Lagrangian Derivation Table~1).}
\end{table}

\section{Relation to relativistic completion}

The relativistic Skyrme-like lift~\cite{rel} is recovered by promoting~\eqref{eq:free-energy}
to Lorentz-invariant form~\eqref{eq:sigma}--\eqref{eq:burst-S}. Observer synchronization
and one-loop remarks of that paper are unchanged. Phase~1.1 adds the product-group
Yang--Mills scaffold and an explicit braiding pin without reopening locks.

\section{Conclusion}

The conduit PDE retains its free-energy / wave-map foundation. Phase~1.1 freezes a
unified action~\eqref{eq:unified-action} with SU(3)$\times$SU(2)$\times$U(1) kinetics,
Hopf topology, and holonomy/braiding terms compatible with $W_g$, $\kappa$, and
$\phi_b$. Gate~A-P blocks unphysical ghosts and lock drift. SM particle content and
emergent gravity remain gated for Phases~2--3.

\begin{thebibliography}{9}
\bibitem{lag}
A.~Kinder, \emph{Derive the Lagrangian for the Conduit PDE}, TOE notes, Apr.~2026.
\bibitem{rel}
A.~Kinder, \emph{Relativistic Completion of the Gauged Hopf Lattice},
invariant\_hunt \texttt{papers/Relativistic\_Completion.tex}, Jul.~2026.
\bibitem{burst}
A.~Kinder, \emph{Burst Threshold Derivation in the Gauged Two-Gyro Hopf Lattice},
invariant\_hunt \texttt{papers/GW\_Burst\_Threshold.tex}, Jun.~2026.
\bibitem{freeze}
invariant\_hunt, \emph{Pre-merger predictive freeze},
\texttt{docs/MILESTONE\_PREMERGER\_PREDICTIVE\_FREEZE.md}, Jul.~2026.
\end{thebibliography}

\end{document}
