\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[hidelinks]{hyperref}

\title{Gauged Hopf Lattice Scaffold Series\\
\large Overview of Papers~I--IV}
\author{Aaron Kinder \\ Independent Researcher}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
This cover note indexes a four-paper scaffold series for the gauged Hopf lattice:
unified action (I), relativistic completion (II), SM content (III), and emergent
gravity (IV). Core locks $W_g=350/\pi$, $\kappa\approx 0.85$, and
$\phi_b^\star\approx 0.8145$ are frozen inputs. Each paper records gated
structure, not a claim of completed SM+GR unification or multi-event GW discovery.
\end{abstract}

\section{Series map}

\begin{center}
\begin{tabular}{@{}clp{0.48\textwidth}@{}}
\hline
\# & Note & Content \\
\hline
I & Unified Action Principle & Free energy, conduit PDE, unified action with
  SU(3)$\times$SU(2)$\times$U(1), Gate~A-P \\
II & Relativistic Completion & Lorentz-invariant lift, quantization outline,
  low-energy targets \\
III & SM Particle Content & Representations, Yukawa scaffold, anomalies, one-loop RG \\
IV & Emergent Gravity & $G_N$ schema, weak-field limit, SI bridge, lattice$\to$metric \\
\hline
\end{tabular}
\end{center}

\section{Shared discipline}

\begin{itemize}
\item Locks $W_g$, $\kappa$, $\phi_b^\star$ are \textbf{not} free fit parameters.
\item FAIL demotes the \emph{mapping} under test, not the locks by default.
\item Independent GW pre-merger prediction pipelines remain score-only and closed
  for bulk PE-power mapping stretch (separate campaign documentation).
\item Software companions, where cited, are optional reproducibility aids.
\end{itemize}

\section{Recommended reading order}

Paper~I $\to$ II for the action; Paper~III for spectrum scaffolding; Paper~IV for
gravity scaffolding. Cross-citations treat companions as peer notes pending arXiv IDs.

\section{Source packaging}

Individual \texttt{.tex} sources and a combined series tarball are produced by
\begin{verbatim}
cd papers && make package
\end{verbatim}
See \texttt{README\_ARXIV.md} for arXiv upload conventions.

\end{document}
