\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[hidelinks]{hyperref}

\title{SM Particle Content from Gauged Hopf Lattice Modes\\
\large Representations, Yukawa Scaffold, Anomalies, and One-Loop Gauge RG}
\author{Aaron Kinder \\ Independent Researcher}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
We record a gated map from excitations of the gauged Hopf lattice
(product Yang--Mills + Hopfions + holonomy/braiding) onto the Standard Model
field content under $G=\mathrm{SU}(3)_c\times\mathrm{SU}(2)_L\times\mathrm{U}(1)_Y$.
Gate~SM-1 requires correct representations and electric charges $Q=T_3+Y/2$.
Three generation \emph{copies} are listed via braiding-layer index.
We also record a topological Yukawa mass ansatz (Gate~SM-2 mass structure),
standard anomaly cancellation, and one-loop SM gauge RG health (Gate~SM-3).
Core locks $W_g=350/\pi$, $\kappa\approx 0.85$, $\phi_b^\star\approx 0.8145$ remain
frozen. No gravity or full mass-spectrum discovery claim is made.
\end{abstract}

\section{Discipline and locks}

Frozen inputs:
\[
W_g=\frac{350}{\pi},\qquad \kappa\approx 0.85,\qquad \phi_b^\star\approx 0.8145.
\]
Unified action~\cite{lag,rel}:
\begin{equation}
S=S_\sigma+S_{\mathrm{Skyrme}}+S_{\mathrm{YM}}
+S_{\mathrm{Hopf}}+S_{\mathrm{hol}}+S_{\mathrm{drive}}
+S_{\mathrm{burst}}+S_{\mathrm{grav}}.
\label{eq:unified-action}
\end{equation}
Pre-merger GW freeze is independent and untouched.

\section{Hypercharge convention}

\begin{equation}
Q = T_3 + \frac{Y}{2}.
\label{eq:Q}
\end{equation}
Assignments (PDG weak hypercharge):
\begin{center}
\begin{tabular}{@{}ll@{}}
\hline
Multiplet & $Y$ \\
\hline
$Q_L$ & $1/3$ \\
$u_R$ & $4/3$ \\
$d_R$ & $-2/3$ \\
$L_L$ & $-1$ \\
$e_R$ & $-2$ \\
$H$ & $1$ \\
$g,W,B$ & $0$ \\
\hline
\end{tabular}
\end{center}

\section{Gauge bosons (adjoint / $U(1)$ modes)}

\begin{center}
\begin{tabular}{@{}llll@{}}
\hline
Field & $({\rm SU}(3),{\rm SU}(2),Y)$ & Lattice mode & Notes \\
\hline
$g$ & $(8,1,0)$ & YM adjoint SU(3) flux flywheel & 8 gluons \\
$W$ & $(1,3,0)$ & YM adjoint SU(2) holonomy wave & $W^\pm,W^3$ pre-EWSB \\
$B$ & $(1,1,0)$ & $U(1)$ global pointer seed & mixes $\to\gamma,Z$ \\
\hline
\end{tabular}
\end{center}

\section{Higgs}

\begin{center}
\begin{tabular}{@{}llll@{}}
\hline
Field & $({\rm SU}(3),{\rm SU}(2),Y)$ & Lattice mode & Charges \\
\hline
$H$ & $(1,2,1)$ & Holonomy modulus / pointer VEV & $H^+$ ($Q=1$), $H^0$ ($Q=0$) \\
\hline
\end{tabular}
\end{center}

\section{Fermions (one generation)}

Hopfion solitons inherit spin $1/2$ from the double cover $S^3\to\mathrm{SO}(3)$.

\begin{center}
\begin{tabular}{@{}llll@{}}
\hline
Field & $({\rm SU}(3),{\rm SU}(2),Y)$ & Chirality & Lattice mode stem \\
\hline
$Q_L$ & $(3,2,1/3)$ & $L$ & \texttt{hopfion\_color\_weak\_doublet} \\
$u_R$ & $(3,1,4/3)$ & $R$ & \texttt{hopfion\_up\_singlet} \\
$d_R$ & $(3,1,-2/3)$ & $R$ & \texttt{hopfion\_down\_singlet} \\
$L_L$ & $(1,2,-1)$ & $L$ & \texttt{hopfion\_lepton\_doublet} \\
$e_R$ & $(1,1,-2)$ & $R$ & \texttt{hopfion\_charged\_lepton\_singlet} \\
\hline
\end{tabular}
\end{center}

\section{Three generations (structure only)}

Generation index $n_g\in\{1,2,3\}$ labels braiding layers / Hopf winding families
pinned by $\phi_b^\star$. Gate~SM-1 requires three \emph{identical} representation
copies.

\section{Phase~2.2: Topological Yukawa ansatz}

Braiding-layer mass hierarchy ($i=0,1,2$ light$\to$heavy, $s\in\{u,d,e\}$):
\begin{equation}
m_i^{(s)}
=\frac{v}{\sqrt{2}}\, y_s\,
\exp\!\bigl(-\beta_s (2-i)^{p_s}\bigr)\,
\bigl(1+\varepsilon_s\cos(\phi_b^\star+\psi_s i)\bigr),
\label{eq:yukawa-mass}
\end{equation}
with $v\simeq 246.22\,\mathrm{GeV}$. Free sector params $\{y_s,\beta_s,p_s,\varepsilon_s,\psi_s\}$;
locks $W_g,\kappa,\phi_b^\star$ frozen (mild dressing only).

CKM via Wolfenstein $(\lambda,A,\bar\rho,\bar\eta)$ with mild $\phi_b^\star$ phase dressing.
A full $Y_{ij}$ braiding matrix is retained for documentation.

Gate~SM-2 mass: Optuna vs PDG $\ln m$ and $|V_{ij}|$ $\chi^2$.
Loose thresholds: $\chi^2_{\ln m}/\mathrm{dof}\le 9$,
$\chi^2_{|V|}/\mathrm{dof}\le 100$. FAIL demotes the \emph{ansatz}, not the locks.
Executable: \texttt{scripts/sm\_yukawa\_ansatz.py --sweep}.

\section{Anomaly cancellation (Gate~SM-3)}

Per generation, with right-handed fields rewritten as left-handed conjugates, the
standard SM anomaly coefficients vanish:
\[
[{\rm SU}(3)]^2{\rm U}(1),\quad
[{\rm SU}(2)]^2{\rm U}(1),\quad
[{\rm U}(1)]^3,\quad
{\rm grav}\mbox{-}{\rm U}(1)
\;\;\to\;\; 0.
\]
Executable: \texttt{src/sm\_mapping.py::check\_anomaly\_cancellation}.

\section{Phase~2.3: One-loop gauge RG}

GUT-normalized couplings $\alpha_i=g_i^2/(4\pi)$ with one-loop running
\begin{equation}
\frac{\mathrm{d}\,\alpha_i^{-1}}{\mathrm{d}\ln\mu}
= -\frac{b_i}{2\pi},
\qquad
b=\Bigl(\tfrac{41}{10},\,-\tfrac{19}{6},\,-7\Bigr)
\quad (n_g=3,\,n_H=1).
\label{eq:rg}
\end{equation}
Boundary data at $M_Z$: $\alpha_{\mathrm{em}}$, $\sin^2\theta_W$, $\alpha_s$.
Gate~SM-3 requires asymptotic freedom of $\alpha_s$, Landau-free evolution on
$[M_Z,10^{16}\,\mathrm{GeV}]$ (one-loop SM), and round-trip consistency.
\textbf{No unification claim.} Executable: \texttt{scripts/sm\_rg\_flow.py}.

\section{Gates}

\begin{center}
\begin{tabular}{@{}llp{0.42\textwidth}@{}}
\hline
Gate & Status & Requirement \\
\hline
SM-1 & PASS (scaffold) & Representations + $Q=T_3+Y/2$ + mode coverage \\
SM-2 & PASS (mass structure) & Hierarchical Yukawa $\chi^2$ under frozen locks \\
SM-3 & PASS (scaffold) & Anomalies + one-loop SM RG health (no unification claim) \\
\hline
\end{tabular}
\end{center}

Companion software (optional reproducibility) implements the same gates.

\section{What is not claimed}

Derived soliton masses from first principles, full CKM/PMNS prediction,
$\alpha_s$ unification, or Einstein gravity. Demote any map that fails multi-gate
consistency or reopens core locks.

\paragraph{Series note.}
This note is Paper~III of the gated scaffold series (Papers~I--IV).

\begin{thebibliography}{9}
\bibitem{lag}
A.~Kinder, \emph{Unified Action Principle for the Gauged Hopf Lattice},
companion note (Paper~I), Jul.~2026.
\bibitem{rel}
A.~Kinder, \emph{Relativistic Completion of the Gauged Hopf Lattice},
companion note (Paper~II), Jul.~2026.
\bibitem{grav}
A.~Kinder, \emph{Emergent Gravity from the Gauged Hopf Lattice},
companion note (Paper~IV), Jul.~2026.
\end{thebibliography}

\end{document}
